---
title: "Bags_analysis"
author: "EVS"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, include=FALSE}
#IMPORTANT INFO:

#eval: evaluate code chunk (chunk won't be run if eval = FALSE)
#echo: include code along with output when kitting file (figures but not code will appear when echo = FALSE)
#include: include output when knitting file 
#cache: save results for future renders (if cache = TRUE and no edits are made to code chunk,
        #it will give previously generated output rather than evaluating again -- 
        #useful for simulations which take a long time to run --
        #I've had some issues with cache not working when I clear the environment or install new packages
        #so I'd recommend saving simulation outputs separately as a plan B if cache fails)


#Sections are designated by headings "#_"
#Some sections have subsections designated by sub headings "##_"
#at the top of each section is an informational code chunk describing the purpose, inputs, and outputs of the section
#most sections correspond to a specific statistical analysis or figure

```

```{r, echo = FALSE, include=FALSE, eval = FALSE}
#clear environment -- only run to clear environment!
rm(list = ls())
```

```{r, echo = FALSE, include=FALSE}
#Packages
install.packages("tidyr")
install.packages("readxl")
install.packages("plyr")
install.packages("dplyr")
install.packages("modeest")
install.packages("ggplot2")
install.packages("vegan")
install.packages("ggpubr")
install.packages("lme4")
install.packages("effects")
install.packages("tidyverse")
install.packages("tibble")
install.packages("lattice")
install.packages("sjPlot")
install.packages("sjmisc")
install.packages("Matrix")
install.packages("nlme")
install.packages("multcomp")
install.packages("emmeans")
install.packages("meta")
install.packages("betapart")
install.packages("xtable")
install.packages("ggpubr")

library(tidyr)
library(readxl)
library(plyr)
library(dplyr)
library(modeest)
library(ggplot2)
library(vegan)
library(ggpubr)
library(lme4)
library(effects)
library(tidyverse)
library(tibble)
library(lattice)
library(sjPlot)
library(sjmisc)
library(Matrix)
library(nlme)
library(multcomp)
library(emmeans)
library(meta)
library(betapart)
library(xtable)
library(ggpubr)

```

# Imports
```{r, echo = FALSE, include=FALSE}
#read in DATA_Combined: master data sheet including Pre, Post1, Post2 by plant
DATA_Combined <- read_excel("DATA_Combined.xlsx")

#create data frame containing only Post1 data
  #Post1 data contains data from both Post surveys in 2019
Post1 <- subset(DATA_Combined, Survey == "Post1")

```

```{r, echo = FALSE, include=FALSE}
#input: DATA_Combined

#output: plants (a data frame showing total abundance of arthropod species found on each plant species 
#across the entire experiment after treatment implementation)

#used as a table in supplement

plants <- DATA_Combined %>%
  filter(Survey == "Post1") %>%
  dplyr::select(-c(Survey, Bag, Block, Year, Treatment, CommonSp, PlantNumber)) %>%
  pivot_longer(2:207, names_to = "Species", values_to = "abundance")%>% 
  group_by(Plant, Species) %>%
  summarise_all(sum) %>%
  pivot_wider(names_from = "Plant", values_from = "abundance")

```

```{r, echo = FALSE, include=FALSE}
#input: DATA_Combined

#output: plants3 (a data frame showing the number of plants of each species in each bag
#across the entire experiment before tretament implementation)

#used as a table in supplement

plants3 <- DATA_Combined %>%
  filter(Survey == "Pre") %>%
  dplyr::select(c(Bag, Plant,Treatment)) %>%
  group_by_all() %>% 
  summarise(Count = n())%>%
  pivot_wider(names_from = "Plant", values_from = "Count") %>%
  replace(is.na(.), 0)

plants3[plants3 == 'C'] <- 'Control'
plants3[plants3 == 'R'] <- 'Random Loss'
plants3[plants3 == 'H'] <- 'Homogenization'

```

```{r, echo = FALSE, include=FALSE}
#input: DATA_Combined

#output: plants3 (a data frame showing the number of plants of each species in each bag
#across the entire experiment

#used as a table in supplement

plants2 <- DATA_Combined %>%
  filter(Survey == "Post1") %>%
  dplyr::select(c(Bag, Plant,Treatment)) %>%
  group_by_all() %>% 
  summarise(Count = n())%>%
  pivot_wider(names_from = "Plant", values_from = "Count") %>%
  replace(is.na(.), 0)

plants2[plants2 == 'C'] <- 'Control'
plants2[plants2 == 'R'] <- 'Random Loss'
plants2[plants2 == 'H'] <- 'Homogenization'

```


# Calculate Speices Richness
```{r, echo = FALSE, include=FALSE}
#condense master data sheet from plant level to bag level
#calculate species richness by bag for Post1 data
#use anova to analyze how species richness is affected by treatment and common species
#plot species richness by common species

#input: Post1, df containing Post1 data by plant
      #to generate, run Imports section

#output: Post1_bag, df containing number of arthropod individuals of each species and total species richness in each bag
        #species richness is the last column
        #plot of species richness by common species
```


```{r, echo = FALSE, include=FALSE}
#DATA MANIPULATION

#use group_by and summarize to condense data frame by bag
#grouping by Bag, Block, Treatment, and Common Species allows us to keep those variables
  #and just take the sum of arthropods in each bag for each arthropod species
Post1_bag <- Post1[,c(2,3,5,8:214)] %>%
  group_by(Bag, Block, Treatment, CommonSp) %>%
  summarize_all(sum)

#calculate species richness for each bag by counting number of arthropod species with >= 1 individual present
#not totally sure how apply works but it does what I need it to
Post1_bag$SpRichness <- apply(Post1_bag[,5:210], MARGIN = 1, function(x)sum(x>=1)) 

```

# Evenness and Effective Number of Species

```{r, echo = FALSE, include=FALSE}
#calculate and visualize diversity indexes: ENS shannon, ENS simpson, evenness, species richness

#input: Post1_bag, df containing Post1 data by bag
      #to generate, run Species Richness by Treatment section

#output: Diversity_df, data frame containing ENS shannon, ENS simpson, Species richness, and evenness for each bag
          #boxplot of ENS shannon by common species
          #boxplot of ENs shannon by treatment
          #line and dot plot of all diversity indeces by treatment
          #line and dot plot of al diversity indeces by treatment
```

```{r, echo = FALSE, include=TRUE}
#DATA ANALYSIS
#ENS simpson

#empty vector for simpson values
simpsonvec <- c()
#for numbers 1 through 45 (each row in Post1_bag)
for(i in 1:45) {
  #calculate the simpson index of the bag based on abundance data
  simpson <- diversity(Post1_bag[i,5:210], "simpson")
  simpson <- 1/(1-simpson)
  #add the simpson value to vector
  simpsonvec <- c(simpsonvec, simpson)
}


```


```{r, echo = FALSE, include=TRUE}
#DATA ANALYSIS
#ENS shannon

#empty vector for shannon values
shannonvec <- c()
#for numbers 1 through 45 (each row in Post1_bag)
for(i in 1:45) {
  #calculate the shannon index of the bag based on abundance data
  shannon <- diversity(Post1_bag[i,5:210], "shannon")
  shannon <- exp(shannon)
  #add the shannon value to vector
  shannonvec <- c(shannonvec, shannon)
}

```

```{r, echo = FALSE, include=TRUE}
#DATA ANALYSIS
#evenness (based on shannon index)

#empty vector for evenness values
evn_vec <- c()
#for numbers 1 through 45 (each row in Post1_bag)
for(i in 1:45) {
  #calculate the evenness of the bag based on abundance data
  shannon_e <- diversity(Post1_bag[i,5:210], "shannon")
  evn <- shannon_e/log(Post1_bag[i,211])
  #add the evenness value to vector
  evn_vec <- c(evn_vec, evn)
}
```

```{r, echo = FALSE, include=FALSE}
#create data frame by bag with Bag, Treatment, Common Species, Species Richness, ENS simpson, ENS Shannon, and evenness columns
#use vectors of diversity indexes created in previous chunks
Diversity_df <- Post1_bag[,c(1,2,3,4,211)]
Diversity_df$Enssimpson <- simpsonvec
Diversity_df$Ensshannon <- shannonvec
Diversity_df$Evenness <- as.numeric(evn_vec)
```

# Plot Species Abundance Distributions
```{r, echo = FALSE, include=TRUE}
#create arthropod species abundance distributions for each common species region and treatment (averaged across bags)

#input: Post1_bag
#output: abundance distribution plots for each common species region and treatment (abundance for each arthropod species averaged at bag level


```

```{r, echo = FALSE, include=TRUE}
#DATA ANALYSIS
#create data frame with average abundances for each arthropod species in each treatment

Treat_abundance <- Post1_bag[,c(3,5:210)] %>%
  #take the mean of arthropod abundances by bag in each treatment
  group_by(Treatment) %>%
  summarise_all(mean) %>%
  #pivot data frame so that arthropod species names are rows
  pivot_longer(cols = 2:207, names_to = "Arthropods", values_to = "Abundance") %>%
  #remove any abundances that are 0
  subset(Abundance > 0) %>%
  #arrange data frame in descending order of arthropod abundance
  arrange(desc(Abundance)) %>%
  #row_number indicates descending order of abundances within each treatment
  #*note: no idea what x = 0 does here but code doesn't work without it 
  group_by(Treatment) %>%
  mutate(rows = row_number(),x = 0)


```

```{r, echo = FALSE, include=TRUE}
#DATA ANALYSIS
#create data frame with average abundances for each arthropod species in each common species region

Comm_abundance <- Post1_bag[,c(4:210)] %>%
  #take the mean of arthropod abundances by bag in each common species region
  group_by(CommonSp) %>%
  summarise_all(mean) %>%
  #pivot data frame so that arthropod species names are rows
  pivot_longer(cols = 2:207, names_to = "Arthropods", values_to = "Abundance") %>%
  #remove any abundances that are 0
  subset(Abundance > 0) %>%
  #arrange data frame in descending order of arthropod abundance
  arrange(desc(Abundance)) %>%
  #row_number indicates descending order of abundances within each common species
  #*note: no idea what x = 0 does here but code doesn't work without it 
  group_by(CommonSp) %>%
  mutate(rows = row_number(), x = 0)
```

```{r, echo = FALSE, include=TRUE}
#DATA ANALYSIS
#this was initially for drawing vertical lines at the abundance tails of the plots, but i realized it was easier to just do it in ppt so it's not being used in the plot

H_treat <- subset(Treat_abundance, Treatment == "H")
vlineH <- max(H_treat$rows)

R_treat <- subset(Treat_abundance, Treatment == "R")
vlineR <- max(R_treat$rows)

C_treat <- subset(Treat_abundance, Treatment == "C")
vlineC <- max(C_treat$rows)

```

```{r, echo = FALSE, include=FALSE}
#DATA ANALYSIS
#this was initially for drawing vertical lines at the abundance tails of the plots, but i realized it was easier to just do it in ppt so it's not being used in the plot

H_comm <- subset(Comm_abundance, CommonSp == "C")
vlineC_comm <- max(H_comm$rows)

R_comm <- subset(Comm_abundance, CommonSp == "B")
vlineB_comm <- max(R_comm$rows)

C_comm <- subset(Comm_abundance, CommonSp == "A")
vlineA_comm <- max(C_comm$rows)

```

```{r, echo = FALSE, include=TRUE}
#PLOT
#arthropod species abundance distribution for each treatment averaged at bag level

#revalue treatment names to how they will show up on plot
Treat_abundance$Treatment <- revalue(Treat_abundance$Treatment, c("C" = "Control", "R" = "Random Loss", "H" = "Homogenization"))
#specify order treatment levels will appear on plot
Treat_abundance$Treatment <- factor(Treat_abundance$Treatment, levels=c("Control", "Random Loss", "Homogenization"))

Treat_abundance %>%
  ggplot() +
  #create bar plot of rows by arthropod abundance
  geom_col(aes(x = rows, y = Abundance, group = rows),
           fill = "black", color = "black", size = .2) +
  #facet by treatment
  facet_wrap(~Treatment, ncol = 1) +
  #specify theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 25, face = "bold"),
        aspect.ratio = 0.35,
        strip.text.x = element_text(color = "black", size = 18, face = "bold", hjust = 0.25),
        axis.text = element_text(size = 14, face = "bold"),
        axis.line = element_line(size = 1)) +
  #change length of y axis
  scale_y_continuous(limits=c(-100, 750)) +
  #customize y axis label
  labs(x = "Arthropod Species") 

Treat_abundance %>%
  ggplot(aes(x = rows, y = Abundance)) +
  #create line plot of rows by arthropod abundance
  geom_line(size= 1.5) +
  #facet by treatment group
  facet_wrap(~Treatment, ncol = 1) +
  #specify theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 25, face = "bold"),
        aspect.ratio = 0.35,
        strip.text.x = element_text(color = "black", size = 18, face = "bold", hjust = 0.25),
        axis.text = element_text(size = 14, face = "bold"),
        axis.line = element_line(size = 1)) +
  #change length of y axis
  scale_y_continuous(limits=c(-50, 500)) +
  #customize y axis lable
  labs(x = "Arthropod Species")
```

```{r, echo = FALSE, include=TRUE}
#PLOT
#arthropod species abundance distribution for each common species region averaged at bag level

#revalue common species names to how they will show up on plot
Comm_abundance$CommonSp <- revalue(Comm_abundance$CommonSp, c("A" = "A. tuberosa", "B" = "B. alba", "C" = "S. altissima"))
#specify order common species levels will appear on plot
Comm_abundance$CommonSp <- factor(Comm_abundance$CommonSp, levels=c("A. tuberosa", "S. altissima", "B. alba"))

Comm_abundance %>%
  ggplot() +
  #create bar plot of rows by arthropod abundance
  geom_col(aes(x = rows, y = Abundance, group = rows),
           fill = "black", color = "black", size = .2) +
  #facet by common species
  facet_wrap(~CommonSp, ncol = 1) +
  #specify theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 25, face = "bold"),
        aspect.ratio = 0.35,
        strip.text.x = element_text(color = "black", size = 18, face = "bold.italic", hjust = 0.25),
        axis.text = element_text(size = 14, face = "bold"),
        axis.line = element_line(size = 1)) +
  #change length of y axis
  scale_y_continuous(limits=c(-50, 500)) +
  #customize y axis lable
  labs(x = "Arthropod Species")


Comm_abundance %>%
  ggplot(aes(x = rows, y = Abundance)) +
  #create line plot of rows by arthropod abundance
  geom_line(size= 1.5) +
  #facet by common species
  facet_wrap(~CommonSp, ncol = 1) +
  #specify theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 25, face = "bold"),
        aspect.ratio = 0.35,
        strip.text.x = element_text(color = "black", size = 18, face = "bold.italic", hjust = 0.25),
        axis.text = element_text(size = 14, face = "bold"),
        axis.line = element_line(size = 1)) +
  #change length of y axis
  scale_y_continuous(limits=c(-50, 500)) +
  #customize y axis lable
  labs(x = "Arthropod Species")

```

# Mixed Model Selection Species Richness
```{r, echo = TRUE, include=TRUE}
#find the model best explaining variation in species richness between bags

#input: Diversity_df
#output: best fit model, pairwise comparisons of fixed effect levels in that model, forest plot showing relative effects of fixed effect levels in best fit model
```
## Choosing Optimal variance Structure
```{r, echo = TRUE, include=TRUE}
#STATISTICS
#determine whether any fixed effects in the species richness model require weighted variances (have non-normal distributions)

#create weights for each fixed effect that could possibly explain species richness (including their interaction)
r1 <- varIdent(form = ~1|Treatment)
r2 <- varIdent(form = ~1|CommonSp)
r3 <- varIdent(form = ~1|Treatment*CommonSp)

#full model containing all fixed effects
rich_gls <- gls(SpRichness ~ Treatment* CommonSp, method = "REML", data = Diversity_df)

#full model plus each respective weighted variance
rich1_gls <- gls(SpRichness ~ Treatment* CommonSp, method = "REML", data = Diversity_df, weights = r1)
rich2_gls <- gls(SpRichness ~ Treatment* CommonSp, method = "REML", data = Diversity_df, weights = r2)
rich3_gls <- gls(SpRichness ~ Treatment* CommonSp, method = "REML", data = Diversity_df, weights = r3)

#compare AIC of full model to each weighted model (no weights required)
anova(rich_gls, rich1_gls, rich2_gls, rich3_gls)

#*note: this weighted variances protocol is based on methods described in Zuur 2009, which said to test weighted models against full model, then include any necessary weights during model selection later. Zuur emphasizes the importance of using REML when selecting weighted variances, not sure why, but it doesn't change the results.

#plot residuals vs. fitted values and residuals vs. fixed effect levels
plot(as.factor(Diversity_df$Treatment), resid(rich_gls))
plot(as.factor(Diversity_df$CommonSp), resid(rich_gls))
plot(rich_gls)


```
\newpage
## model selection
```{r, echo = TRUE, include=TRUE}
#STATISTICS
#select best fit model explaining variation in species richness

#full model containing all fixed effects and their interactions plus random effect (block)
rich1 = lme(SpRichness ~ Treatment + CommonSp + Treatment*CommonSp , method = "ML", random = ~1|Block, data = Diversity_df)
#drop interaction term
rich2 = lme(SpRichness ~ Treatment + CommonSp, method = "ML", random = ~1|Block, data = Diversity_df)
#drop common species term
rich3 = lme(SpRichness ~ Treatment, method = "ML", random = ~1|Block, data = Diversity_df)
#drop richness term
rich4 = lme(SpRichness ~ CommonSp, method = "ML", random = ~1|Block, data = Diversity_df)

#select best fit model based on AIC values (best fit is rich 2)
anova(rich1, rich2, rich3, rich4)
anova(rich1, rich2)
anova(rich2, rich3)

#diagnostic plots: checking for normally distributed residuals
qqnorm(resid(rich2))
qqline(resid(rich2))

plot(fitted(rich2),resid(rich2, type = "normalized"), abline(h= 0))

#pairwise comparisons for common species levels
emms_rich2 <- emmeans(rich2, ~CommonSp)
summary(contrast(emms_rich2, interaction = "pairwise"))
print(emms_rich2)

#pairwise comparisons for common species levels
emms_rich3 <- emmeans(rich2, ~Treatment)
summary(contrast(emms_rich3, interaction = "pairwise"))
print(emms_rich3)

emms_rich2_re <- as.data.frame(emms_rich2)
emms_rich2_re$CommonSp <- revalue(emms_rich2_re$CommonSp, c("C" = "S. altissima", "B" =  "B. alba", "A" = "A. tuberosa"))
emms_rich2_re$CommonSp <- factor(emms_rich2_re$CommonSp, levels=c("S. altissima", "A. tuberosa", "B. alba"))

#plot averages of all 3 diversity indices grouped for each Common Species
ggplot(data = emms_rich2_re, aes(x= CommonSp, y=emmean, group = CommonSp)) +
  #add error bars based on sd values in data frame
  geom_errorbar(aes(ymin = emmean- SE, ymax = emmean+SE), width = 0.2)+
  #add and customize lines and points (2 layers points with one slightly larger create outlines)
  #geom_line(aes(color = CommonSp), lwd = 2)+
  geom_point(size = 5) +
  ylab("Species Richness") + 
  #add theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 22, face = "bold"),
        axis.title.x=element_blank(),
        aspect.ratio = 1,
        axis.text.x = element_text(size = 22, face = "bold", color = "black", angle = 50, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 22, face = "bold", color = "black"),
        axis.line = element_line(size = 1))


emms_rich_re <- as.data.frame(emms_rich)
emms_rich_re$Treatment <- revalue(emms_rich_re$Treatment, c("H" = "Habitat Simplification", "R" =  "Habitat Loss", "C" = "Control"))
emms_rich_re$Treatment <- factor(emms_rich_re$Treatment, levels=c("Control", "Habitat Loss", "Habitat Simplification"))

emms_rich_re$CommonSp <- revalue(emms_rich_re$CommonSp, c("C" = "S. altissima", "B" =  "B. alba", "A" = "A. tuberosa"))
emms_rich_re$CommonSp <- factor(emms_rich_re$CommonSp, levels=c("S. altissima", "A. tuberosa", "B. alba"))

#plot averages of all 3 diversity indices grouped for each Common Species
ggplot(data = emms_rich_re, aes(x= Treatment, y=emmean, group = Treatment)) +
  #add error bars based on sd values in data frame
  geom_errorbar(aes(ymin = emmean-SE, ymax = emmean+SE), width = 0.2)+
  #add and customize lines and points (2 layers points with one slightly larger create outlines)
  geom_point(size = 5) +
  ylab("Species Richness") + 
  #add theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 22, face = "bold"),
        axis.title.x=element_blank(),
        aspect.ratio = 1,
        axis.text.x = element_text(size = 22, face = "bold", color = "black", angle = 50, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 22, face = "bold", color = "black"),
        axis.line = element_line(size = 1))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) 


#plot averages of all 3 diversity indices grouped for each Common Species
ggplot(data = emms_rich_re, aes(x= CommonSp, y=emmean, group = Treatment), position = position_dodge(.5)) +
  #add error bars based on sd values in data frame
  geom_errorbar(aes(ymin = emmean-SE, ymax = emmean+SE), width = 0.2, position = position_dodge(.5))+
  #add and customize lines and points (2 layers points with one slightly larger create outlines)
  #geom_line(aes(color = CommonSp), lwd = 2)+
  geom_point(size = 5, aes(shape = factor(Treatment)), position = position_dodge(.5)) +
  geom_point(size = 3, aes(shape = factor(Treatment)), position = position_dodge(.5), color = "white") +
  ylab("Species Richness") + 
  #add theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 22, face = "bold"),
        axis.title.x=element_blank(),
        aspect.ratio = 1,
        axis.text.x = element_text(size = 22, face = "bold", color = "black", angle = 50, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 22, face = "bold", color = "black"),
        axis.line = element_line(size = 1),
        legend.title = element_blank(),
        legend.text = element_text(size = 22, face = "bold", color = "black"))
  


```

# Mixed Model Selection Ensshannon
## Choosing optimal variance structure
```{r, echo = TRUE, include=TRUE}
#STATISTICS
#determine whether any fixed effects in the ENS shannon model require weighted variances (have non-normal distributions)

#full model containing all fixed effects
shan_gls <- gls(Ensshannon ~ Treatment* CommonSp, method = "REML", data = Diversity_df)

#full model plus each respective weighted variance
shan1_gls <- gls(Ensshannon ~ Treatment* CommonSp, method = "REML", data = Diversity_df, weights = r1)
shan2_gls <- gls(Ensshannon ~ Treatment* CommonSp, method = "REML", data = Diversity_df, weights = r2)
shan3_gls <- gls(Ensshannon ~ Treatment* CommonSp, method = "REML", data = Diversity_df, weights = r3)

#compare AIC of full model to each weighted model 
#weights required for treatment and common species effects, so optimal variance structure includes r3
anova(shan_gls, shan1_gls, shan2_gls, shan3_gls)

#plot residuals vs. fitted values and residuals vs. fixed effect levels
plot(as.factor(Diversity_df$Treatment), resid(shan3_gls, type = "normalized"))
plot(as.factor(Diversity_df$CommonSp), resid(shan3_gls, type = "normalized"))
plot(shan3_gls)
```
\newpage

## Model selection
```{r, echo = TRUE, include=TRUE}
#STATISTICS
#select best fit model explaining variation in ENS shannon

#full model containing all fixed effects and their interactions plus random effect (block)
shan1 = lme(Ensshannon ~ Treatment + CommonSp + Treatment*CommonSp , weights = r3, method = "ML", random = ~1|Block, data = Diversity_df)
#drop interaction term
shan2 = lme(Ensshannon ~ Treatment + CommonSp, weights = r3, method = "ML", random = ~1|Block, data = Diversity_df)
#drop Common Species term
shan3 = lme(Ensshannon ~ Treatment, weights = r3, method = "ML", random = ~1|Block, data = Diversity_df)
#drop treatment term
shan4 = lme(Ensshannon ~ CommonSp,  weights = r3, method = "ML", random = ~1|Block, data = Diversity_df)

#select best fit model based on AIC values 
#AIC shan1 = AIC shan4, so choose shan4 because it has the fewest terms
anova(shan1, shan2, shan3, shan4)
anova(shan1, shan4)

#diagnostic plots: checking for normally distributed residuals
qqnorm(resid(shan1))
qqline(resid(shan1))

plot(fitted(shan1),resid(shan1, type = "normalized"), abline(h= 0))
  
#pairwise comparisons for common species levels (do not include pairwise comparisons for treatment levels because treatment term was not in best fit model)
emms_shan1 <- emmeans(shan1, ~Treatment*CommonSp)
summary(contrast(emms_shan1, interaction = "pairwise"))
print(emms_shan1)

 emms_shan2 <- emmeans(shan1, ~Treatment)
summary(contrast(emms_shan2, interaction = "pairwise"))
print(emms_shan2)

emms_shan3 <- emmeans(shan1, ~CommonSp)
summary(contrast(emms_shan3, interaction = "pairwise"))
print(emms_shan3)


emms_shan1_re <- as.data.frame(emms_shan1)
emms_shan1_re$CommonSp <- revalue(emms_shan1_re$CommonSp, c("C" = "S. altissima", "B" =  "B. alba", "A" = "A. tuberosa"))
emms_shan1_re$CommonSp <- factor(emms_shan1_re$CommonSp, levels=c("S. altissima", "A. tuberosa", "B. alba"))

emms_shan1_re$Treatment <- factor(emms_shan1_re$Treatment, levels=c("C", "R", "H"))
emms_shan1_re$Treatment <- revalue(emms_shan1_re$Treatment, c("C" = "Control", "H" =  "Simplification", "R" = "Habitat Loss"))


#plot averages of all 3 diversity indices grouped for each Common Species
ggplot(data = emms_shan1_re, aes(x= CommonSp, y=emmean, group = Treatment), position = position_dodge(.5)) +
  #add error bars based on sd values in data frame
  geom_errorbar(aes(ymin = emmean-SE, ymax = emmean+SE), width = 0.2, position = position_dodge(.5))+
  #add and customize lines and points (2 layers points with one slightly larger create outlines)
  #geom_line(aes(color = CommonSp), lwd = 2)+
  geom_point(size = 5, aes(shape = factor(Treatment)), position = position_dodge(.5)) +
  geom_point(size = 3, aes(shape = factor(Treatment)), position = position_dodge(.5), color = "white") +
  ylab("ENS") + 
  #add theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 22, face = "bold"),
        axis.title.x=element_blank(),
        aspect.ratio = 1,
        axis.text.x = element_text(size = 22, face = "bold", color = "black", angle = 50, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 22, face = "bold", color = "black"),
        axis.line = element_line(size = 1),
        legend.title = element_blank(),
        legend.text = element_text(size = 22, face = "bold", color = "black"))

emms_shan2 <- emmeans(shan1, ~Treatment)
summary(contrast(emms_shan2, interaction = "pairwise"))
print(emms_shan2)


emms_shan2_re <- as.data.frame(emms_shan2)

emms_shan2_re$Treatment <- factor(emms_shan2_re$Treatment, levels=c("C", "R", "H"))
emms_shan2_re$Treatment <- revalue(emms_shan2_re$Treatment, c("C" = "Control", "H" =  "Simplification", "R" = "Habitat Loss"))


#plot averages of all 3 diversity indices grouped for each Common Species
ggplot(data = emms_shan2_re, aes(x= Treatment, y=emmean)) +
  #add error bars based on sd values in data frame
  geom_errorbar(aes(ymin = emmean-SE, ymax = emmean+SE), width = 0.2)+
  #add and customize lines and points (2 layers points with one slightly larger create outlines)
  #geom_line(aes(color = CommonSp), lwd = 2)+
  geom_point(size = 5) +
  ylab("ENS") + 
  #add theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 22, face = "bold"),
        axis.title.x=element_blank(),
        aspect.ratio = 1,
        axis.text.x = element_text(size = 22, face = "bold", color = "black", angle = 50, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 22, face = "bold", color = "black"),
        axis.line = element_line(size = 1))

emms_shan3 <- emmeans(shan1, ~CommonSp)
summary(contrast(emms_shan3, interaction = "pairwise"))
print(emms_shan3)


emms_shan3_re <- as.data.frame(emms_shan3)

emms_shan3_re$CommonSp <- revalue(emms_shan3_re$CommonSp, c("C" = "S. altissima", "B" =  "B. alba", "A" = "A. tuberosa"))
emms_shan3_re$CommonSp <- factor(emms_shan3_re$CommonSp, levels=c("S. altissima", "A. tuberosa", "B. alba"))



#plot averages of all 3 diversity indices grouped for each Common Species
ggplot(data = emms_shan3_re, aes(x= CommonSp, y=emmean)) +
  #add error bars based on sd values in data frame
  geom_errorbar(aes(ymin = emmean-SE, ymax = emmean+SE), width = 0.2)+
  #add and customize lines and points (2 layers points with one slightly larger create outlines)
  #geom_line(aes(color = CommonSp), lwd = 2)+
  geom_point(size = 5) +
  ylab("ENS") + 
  #add theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 22, face = "bold"),
        axis.title.x=element_blank(),
        aspect.ratio = 1,
        axis.text.x = element_text(size = 22, face = "bold", color = "black", angle = 50, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 22, face = "bold", color = "black"),
        axis.line = element_line(size = 1))


```
\newpage

# Mixed Model Selection Evenness
## Choosing optimal variance structure
```{r, echo = TRUE, include=TRUE}
#full model containing all fixed effects
evn_gls <- gls(Evenness ~ Treatment* CommonSp, method = "REML", data = Diversity_df)

#full model plus each respective weighted variance
evn1_gls <- gls(Evenness ~ Treatment* CommonSp, method = "REML", data = Diversity_df, weights = r1)
evn2_gls <- gls(Evenness ~ Treatment* CommonSp, method = "REML", data = Diversity_df, weights = r2)
evn3_gls <- gls(Evenness ~ Treatment* CommonSp, method = "REML", data = Diversity_df, weights = r3)

#compare AIC of full model to each weighted model 
#weights required for treatment and common species effects, so optimal variance structure includes r3
anova(evn_gls, evn1_gls, evn2_gls, evn3_gls)

#plot residuals vs. fitted values and residuals vs. fixed effect levels
plot(as.factor(Diversity_df$Treatment), resid(evn_gls, type = "normalized"))
plot(as.factor(Diversity_df$CommonSp), resid(evn_gls, type = "normalized"))
plot(evn_gls)
```
\newpage

## Model selection
```{r, echo = TRUE, include=TRUE}
#STATISTICS
#select best fit model explaining variation in ENS shannon

#full model containing all fixed effects and their interactions plus random effect (block)
evn1 = lme(Evenness ~ Treatment + CommonSp + Treatment*CommonSp, method = "ML", random = ~1|Block, data = Diversity_df)
#drop interaction term
evn2 = lme(Evenness ~ Treatment + CommonSp, method = "ML", random = ~1|Block, data = Diversity_df)
#drop Common Species term
evn3 = lme(Evenness ~ Treatment, method = "ML", random = ~1|Block, data = Diversity_df)
#drop treatment term
evn4 = lme(Evenness ~ CommonSp, method = "ML", random = ~1|Block, data = Diversity_df)

#select best fit model based on AIC values 
#AIC evn1 > AIC evn4
anova(evn1, evn2, evn3, evn4)
anova(evn1, evn4)


#diagnostic plots: checking for normally distributed residuals
qqnorm(resid(evn1))
qqline(resid(evn1))

plot(fitted(evn1),resid(evn1, type = "normalized"), abline(h= 0))

#pairwise comparisons for common species levels (do not include pairwise comparisons for treatment levels because treatment term was not in best fit model)
emms_evn1 <- emmeans(evn1, ~Treatment*CommonSp)
summary(contrast(emms_evn1, interaction = "pairwise"))
print(emms_evn1)

emms_evn2 <- emmeans(evn1, ~CommonSp)
summary(contrast(emms_evn2, interaction = "pairwise"))
print(emms_evn2)

emms_evn3 <- emmeans(evn1, ~Treatment)
summary(contrast(emms_evn3, interaction = "pairwise"))
print(emms_evn3)

emms_evn1_re <- as.data.frame(emms_evn1)
emms_evn1_re$CommonSp <- revalue(emms_evn1_re$CommonSp, c("C" = "S. altissima", "B" =  "B. alba", "A" = "A. tuberosa"))
emms_evn1_re$CommonSp <- factor(emms_evn1_re$CommonSp, levels=c("S. altissima", "A. tuberosa", "B. alba"))

emms_evn1_re$Treatment <- factor(emms_evn1_re$Treatment, levels=c("C", "R", "H"))
emms_evn1_re$Treatment <- revalue(emms_evn1_re$Treatment, c("C" = "Control", "H" =  "Simplification", "R" = "Habitat Loss"))


#plot averages of all 3 diversity indices grouped for each Common Species
ggplot(data = emms_evn1_re, aes(x= CommonSp, y=emmean, group = Treatment), position = position_dodge(.5)) +
  #add error bars based on sd values in data frame
  geom_errorbar(aes(ymin = emmean-SE, ymax = emmean+SE), width = 0.2, position = position_dodge(.5))+
  #add and customize lines and points (2 layers points with one slightly larger create outlines)
  #geom_line(aes(color = CommonSp), lwd = 2)+
  geom_point(size = 5, aes(shape = factor(Treatment)), position = position_dodge(.5)) +
  geom_point(size = 3, aes(shape = factor(Treatment)), position = position_dodge(.5), color = "white") +
  ylab("Evenness") + 
  #add theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 22, face = "bold"),
        axis.title.x=element_blank(),
        aspect.ratio = 1,
        axis.text.x = element_text(size = 22, face = "bold", color = "black", angle = 50, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 22, face = "bold", color = "black"),
        axis.line = element_line(size = 1),
        legend.title = element_blank(),
        legend.text = element_text(size = 22, face = "bold", color = "black"))

emms_evn2 <- emmeans(evn1, ~Treatment)
summary(contrast(emms_evn2, interaction = "pairwise"))
print(emms_evn2)


emms_evn2_re <- as.data.frame(emms_evn2)

emms_evn2_re$Treatment <- factor(emms_evn2_re$Treatment, levels=c("C", "R", "H"))
emms_evn2_re$Treatment <- revalue(emms_evn2_re$Treatment, c("C" = "Control", "H" =  "Simplification", "R" = "Habitat Loss"))


#plot averages of all 3 diversity indices grouped for each Common Species
ggplot(data = emms_evn2_re, aes(x= Treatment, y=emmean)) +
  #add error bars based on sd values in data frame
  geom_errorbar(aes(ymin = emmean-SE, ymax = emmean+SE), width = 0.2)+
  #add and customize lines and points (2 layers points with one slightly larger create outlines)
  #geom_line(aes(color = CommonSp), lwd = 2)+
  geom_point(size = 5) +
  ylab("Evennness") + 
  #add theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 22, face = "bold"),
        axis.title.x=element_blank(),
        aspect.ratio = 1,
        axis.text.x = element_text(size = 22, face = "bold", color = "black", angle = 50, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 22, face = "bold", color = "black"),
        axis.line = element_line(size = 1))

emms_evn3 <- emmeans(evn1, ~CommonSp)
summary(contrast(emms_evn3, interaction = "pairwise"))
print(emms_evn3)


emms_evn3_re <- as.data.frame(emms_evn3)

emms_evn3_re$CommonSp <- revalue(emms_evn3_re$CommonSp, c("C" = "S. altissima", "B" =  "B. alba", "A" = "A. tuberosa"))
emms_evn3_re$CommonSp <- factor(emms_evn3_re$CommonSp, levels=c("S. altissima", "A. tuberosa", "B. alba"))



#plot averages of all 3 diversity indices grouped for each Common Species
ggplot(data = emms_evn3_re, aes(x= CommonSp, y=emmean)) +
  #add error bars based on sd values in data frame
  geom_errorbar(aes(ymin = emmean-SE, ymax = emmean+SE), width = 0.2)+
  #add and customize lines and points (2 layers points with one slightly larger create outlines)
  #geom_line(aes(color = CommonSp), lwd = 2)+
  geom_point(size = 5) +
  ylab("Evenness") + 
  #add theme
  theme_classic() +
  #customize theme
  theme(title = element_text(color = "black", size = 22, face = "bold"),
        axis.title.x=element_blank(),
        aspect.ratio = 1,
        axis.text.x = element_text(size = 22, face = "bold", color = "black", angle = 50, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 22, face = "bold", color = "black"),
        axis.line = element_line(size = 1))


```